Finite-size scaling of free energy in the dimer model on a hexagonal domain

### 2019-12-10 AffineLieAlgebras/MQFT2018/paper.tex
Text on central charge in Introduction is rewritten

Study of dimer models on various lattices and domains led to interesting connections with the
geometry of curved manifolds and with spectra of discrete and continuous Dirac and Laplace operators
\cite{kenyon2002laplacian,kenyon2000asymptotic}. Scaling limit of dimer model is proven to be
described by a Gaussian free-field theory \cite{kenyon2001dominos}, but finite-size corrections were
not considered previously. These corrections are important to close the gap between numerical
simulations and theoretical results.

@@

In the case of a corner singularity on the boundary with the angle
$\Theta$ the logarithmic contribution to the free energy is changed to
\begin{equation}
  \label{eq:86}
  \delta F = \frac{{\bf c}}{24}\left(\frac{\Theta}{\pi}-\frac{\pi}{\Theta}\right) \ln L.
\end{equation}
This formula was recently used by N. Allegra \cite{allegra2015exact} to show that the central charge
in the dimer model on the square lattice with a corner monomer is ${\bf c}=1$, contrary to some
previous claims of ${\bf c}=-2$.

@@

Using Cardy-Peschel formulas, presented above, $\delta F$ can be
computed from value of the central charge for Gaussian free field
${\bf c}=1$, the curvature of metric, induced by the limit shape and
conformal map from the internal region of the ``Arctic curve'' to the
upper half-plane. Such a computation can be used to check our result
on $f_2$ but is quite lengthy and will be presented in a separate
publication.

@@

In the present paper we computed the asymptotic expansion of the free
energy in the dimer model on a hexagonal domain of the hexagonal
lattice, on a hexagon with a side of infinite length and in the model
with coordinate-dependent Boltzmann weights. We presented numerical
results supporting our computations. We've discussed the physical
meaning of the expansion coefficients and argued that our results
support the identification of the scaling behavior of the dimer model
with the Gaussian free field theory.

>>>>

Study of dimer models on various lattices and domains led to interesting connections with the
geometry of curved manifolds and with spectra of discrete and continuous Dirac and Laplace operators
\cite{kenyon2002laplacian,kenyon2000asymptotic}. Scaling limit of dimer model is proven to be
described by a Gaussian free-field theory \cite{kenyon2001dominos}, but finite-size corrections to
the scaling of free energy were previously considered mostly in the absence of the limit shape
\cite{Sh_Izmailian_2019,izmailian2016finite,izmailian2011dimer,izmailian2007non,izmailian2005logarithmic}.

Finite size corrections are important to close the gap between numerical simulations and
theoretical results. Moreover, they can be used to choose the proper field-theoretic description of
the scaling limit. For example, logarithmic term is supposed to be universal and dependent on the
value of central charge ${\bf c}$ of conformal field theory \cite{cardy1988finite}.

Computation of finite-size corrections to the scaling of free energy in dimer model on a rectangular
region on the square lattice led to a series of publications and even controversy about the proper CFT
description of the scaling limit, when some publications claimed ${\bf c}=-2$, while others -- ${\bf c}=1$. 
\cite{allegra2015exact,morin2016integrability,Sh_Izmailian_2019,izmailian2016finite,izmailian2005logarithmic}. 
Note that in this case limit shape does not appear due to the choice of boundary conditions.

The case of limit shape is even more complicated. While the fluctuations the limit shape can be
computed from the variational minimization problem for surface tension and fluctuations are
described by the Gaussian Free Field, the behavior on the boundary of the frozen region is not well
understood. Surface tension becomes singular near the boundary of frozen domain, thus boundary
contribution to the logarithmic correction to free energy scaling cannot be easily obtained from
field-theoretic point of view. Because of this, the computation of finite size corrections from the
exact solution of the model, especially of the logarithmic term, is very interesting.

@@

In the case of a corner singularity on the boundary with the angle
$\Theta$ the logarithmic contribution to the free energy is changed to
\begin{equation}
  \label{eq:86}
  \delta F = \frac{{\bf c}}{24}\left(\frac{\Theta}{\pi}-\frac{\pi}{\Theta}\right) \ln L.
\end{equation}
This formula was recently used by N. Allegra \cite{allegra2015exact} to show that the central charge
in the height-function description of the dimer model on the square lattice with a corner monomer is
${\bf c}=1$, contrary to some previous claims of ${\bf c}=-2$ \cite{morin2016integrability}.

@@

Using Cardy-Peschel formulas, presented above, it should be possible to compute $\delta F$ from
value of the central charge for Gaussian free field ${\bf c}=1$, the curvature of metric, induced by
the limit shape and conformal map from the internal region of the ``Arctic curve'' to the upper
half-plane.
But the metric is singular near the boundary of the internal region and singularities are not
explicitly studied yet \cite{Astala}. Thus our result for value of $f_2$ should be useful in such a
study, which we will attempt to do in a separate publication.

@@

In the present paper we computed the asymptotic expansion of the free energy in the dimer model on a
hexagonal domain of the hexagonal lattice, on a hexagon with a side of infinite length and in the
model with coordinate-dependent Boltzmann weights. We presented numerical results supporting our
computations. We've discussed the physical meaning of the expansion coefficients and argued that our
results support the universal character of the logarithmic correction and can be used to study the
behavior on the boundary of frozen region.
